---
title: "Surveillance Report"
author: "Devin Myers"
output:
  pdf_document:
    toc: true
    highlight: zenburn
---

\vspace{15mm}

\begin{Huge}
\begin{center}

`r format(Sys.Date(), '%B %d, %Y')`  
\end{center}
\end{Huge}

\newpage

Essence Report for `r format(Sys.Date(), '%B %d, %Y')`.

# Introduction 


```{r, echo=FALSE, comment=NA, message=FALSE} 
#libraries
library(ggplot2)
library(dplyr)
library(knitr)
library(ggthemes)
library (rgdal)
library(RColorBrewer)
opts_chunk$set(comment = NA, message = FALSE, echo = FALSE, results = 'hide', fig.height = 3, warnings = FALSE)
```

```{r}
suppressWarnings(library(xtable))
options(xtable.comment = FALSE)

#Get all of our functions
source("functions.R")

#read in the essence data
df<-read.delim("PlainDataDetailsServlet",
               header=TRUE,
               sep=",")
#read in the visits of interest 
voi<- read.delim("voi",
               header=TRUE,
               sep=",")
#read in ili visits
ili<- read.delim("ili",
               header=TRUE,
               sep=",")

#format date 
df$Date <- as.Date(df$Date, format = '%m/%d/%Y')
voi$Date <- as.Date(voi$Date, format = '%m/%d/%Y')
ili$Date <- as.Date(ili$Date, format = '%m/%d/%Y')


# Get today's date
today<- Sys.Date()

# Get yesterday's date
yesterday<- today-1

#Get the day before yestersday date
daybefore_yesterday <- today-2

#Get Alachua Zipcodes spatial data
zip_map <- readOGR("alachuazipcodes", "ACDPS_zipcode")

#Make a new column in zip_map named zip_code
# (make it numeric)
zip_map@data$zip_code <- as.numeric(as.character(zip_map@data$ZIP))

# Keep only real zip codes
zip_map <- zip_map[!is.na(zip_map$zip_code),]



```


This is the ESSENCE surveillance report. There were `r nrow(df)` visits over the past 7 days.

# Visits of Interest  

There were `r nrow(voi)` visits over the past 7 days.  

```{r, results = 'asis'}
good_columns <- c('MedRecNo', 'HospitalName', 'Age', 'Zipcode', 'Sex', 'ChiefComplaintOrig')
sub_voi <- voi[,good_columns]
temp <- xtable(sub_voi)
print(temp, size = '\\small')
```


# Overview  

## Yesterday

There were `r nrow(df[df$Date == yesterday,])` Visits yesterday.

## Sex

Female visits yesterday were `r nrow(df[df$Sex == 'F' & df$Date == yesterday, ])`

Male visits yesterday were `r nrow(df[df$Sex == 'M' & df$Date == yesterday,])`

```{r}
ggplot(data= df, aes (x=Sex))+
  geom_bar(fill= "blue", aplha = 0.6) +
  theme_fivethirtyeight()+
  ggtitle("Male and Female")

```


```{r}
ggplot(data= df, aes (x=Sex)) +
  geom_bar(fill = "green", alpha = 0.8) +
  facet_grid(. ~ Date)+
  theme_fivethirtyeight()+
  ggtitle("Last 7 days")

```

```{r, fig.height= 10}
ggplot(data= df, aes (x=Sex))+
  geom_bar(fill = "turquoise") +
  facet_grid(Age.Group ~ Date) +
  theme_fivethirtyeight()+
  ggtitle("Last 7 days M/F by Age Groups")

```

## Age

Here is the breakdown by age.

```{r,echo=FALSE} 
ggplot(data = df, aes(x=Age))+
  geom_histogram(fill = "salmon", alpha = 0.9) +
  theme_fivethirtyeight() +
  ggtitle('Visits by Age group')
```

## Sex and age


```{r}
ggplot(data= df, aes (x=Sex))+
  geom_bar(fill= "lightblue", alpha = 1) +
  facet_grid(. ~ Age.Group)+
  theme_fivethirtyeight()+
  
    ggtitle("Sex by Age Group")

```


## Geography  

```{r, fig.height=5} 
by_region<- df %>%
  filter(Date == yesterday) %>%
  group_by(Region) %>%
  summarise(visits =n())
by_region <- arrange(by_region, visits)
by_region$Region <- factor(by_region$Region, levels = by_region$Region)
ggplot(data= by_region, aes(x=Region, y=visits))+
  geom_bar(stat="identity", fill = 'springgreen', alpha = 0.9) +
  theme_fivethirtyeight()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle('Visits by County')

```


## By zip code in Alachua only

```{r}
# only take yesterday and only take alachua residents
bza <- df %>% 
  filter(Date == yesterday) %>%
  filter(Region == "Alachua") %>%
  group_by(Zipcode) %>%
    summarise(visits = n()) %>%
  arrange(desc(visits))
bza$Zipcode <- factor(bza$Zipcode, levels = bza$Zipcode)

ggplot(data = bza, aes(x=Zipcode, y = visits))+
  geom_bar(stat="identity", fill= "purple", alpha = 1)+
  theme_fivethirtyeight()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("Alachua County Visits by Zipcode")
```


## Overall number of visits

```{r}
by_date<- df%>%
  group_by(Date)%>%
  summarise (visits =n())
ggplot(data= by_date, aes(x= Date, y=visits)) +
  geom_line(color= "red")+
  geom_smooth(color = "green", fill= "orange")+
  theme_fivethirtyeight()+
  ggtitle("Visits for last 7 Days")


```


# Syndromes  

## ILI  

```{r}
#Create a Data frame for ILI
recent_ili<- df[grepl('ILI', df$Category_flat),]
recent_ili_by_day<- recent_ili %>%
  group_by(Date) %>%
  summarise(visits = n())

ggplot(data = recent_ili_by_day, aes(x= Date, y= visits))+
  geom_bar(stat = "identity", fill = "royalblue")+
  theme_fivethirtyeight()+
  ggtitle("ILI for last 7 days")
```

### Trends  



```{r}
temp <- ili %>%
  group_by(Date) %>%
  summarise(visits = n()) %>% 
  filter(Date >= (today - 30))

ggplot(data = temp, aes(x = Date, y = visits)) +
  geom_point(color = "blue", alpha = 0.2) +
  geom_line(color = "blue", alpha = 0.5) +
  geom_smooth(color = "orange") +
  ggtitle("Last 30 Days")
```


```{r, fig.height= 4.5}
temp <- ili %>%
  group_by(Date) %>%
  summarise(visits = n())

ggplot(data = temp, aes(x = Date, y = visits)) +
  geom_point(color = "blue", alpha = 0.2) +
  geom_line(color = "blue", alpha = 0.5) +
  ggtitle("All Time")
```


### Age  

```{r, warning=FALSE}

#Break ILI up be time 
ili$is_recent <- ifelse(ili$Date >= (today - 30),
                        'Last 30 days',
                        'More than 30 days ago')

# Plot the distribution of age
ggplot(data = ili, aes(x= Age)) +
  geom_density(fill = "blue", color = "orange" ,alpha = 0.5) +
  facet_grid(.~ is_recent, scales= "free_y")


```


## GI

The Bargraph below shows gastrointestinal illnesses within the past 7 days 

```{r}
gi<- df[grepl('GI', df$Category_flat),]
gi_by_day<- gi %>%
  group_by(Date) %>%
  summarise(visits = n())
  
ggplot(data = gi_by_day, aes(x= Date, y= visits))+
  geom_bar(stat = "identity", fill = "orangered")+
  theme_fivethirtyeight()+
  ggtitle("GI last 7 days")

```

```{r, fig.height = 6, warning=FALSE}
# Map
make_map("GI","Reds")
```

## Neuro  

The Bargraph below shows cases of neuro within the past 7 days

```{r}
neuro<- df[grep('Neuro', df$Category_flat),]
neuro_by_day<- neuro %>%
  group_by(Date) %>%
  summarise(visits = n())

ggplot(data = neuro_by_day, aes(x= Date, y= visits))+
  geom_bar(stat = "identity", fill = "limegreen")+
  theme_fivethirtyeight()+
  ggtitle("Neuro Last 7 days")

```

```{r, fig.height = 6}
make_map("Neuro", "Greens")
```

## Rash  

The Bargraph below shows cases of rash within the past 7 days

```{r}
rash<- df[grep('Rash', df$Category_flat),]
rash_by_day<- rash %>%
  group_by(Date) %>%
  summarise(visits = n())

ggplot(data = rash_by_day, aes(x= Date, y= visits))+
 geom_bar(stat = "identity", fill = "red")+ 
 theme_fivethirtyeight()+
  ggtitle("Rash last 7 Days")
```


## Respiratory  

The Bargraph below shows cases of respiratory within the past 7 days 

```{r}
respiratory<- df[grep('Resp', df$Category_flat),]
respiratory_by_day<- respiratory %>%
  group_by(Date) %>%
  summarise(visits = n())

ggplot(data = respiratory_by_day, aes(x= Date, y= visits))+
  geom_bar(stat = "identity", fill = "magenta")+
  theme_fivethirtyeight()+
  ggtitle("Respiratory last 7 days")
```